---
title: "Data wrangling and pre-analysis"
author: "Nikola Gersak"
date: '21 06 2022 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(readxl)
library(fuzzyjoin)

```

# Wrangling: Cadastral parcels (KC) from Odluka merging with cadastral units (KO) and then with cadastral parcel shapes from e

There are three groups of data that are being read into the global environment. First type concerning the 'Odluka' is the data that comes from digitized tables. Second data named 'KO' concerns the cadastral units of whole Croatia, but also their relation to regions and forestry units of Croatia. The third data that is read into the environment 'e' is a database of cadastral parcels of Croatia.

The general objective is to join the cadastral parcels from the Odluka with parcels from the environment 'e' to obtain the variables that can be joined later in the QGIS with it's geometries.

Later also, the statistical analysis will be performed.

## The loading part {-}

```{r}
#read the main file with read.table because read_csv2 doesn't recognise encoding

odluka_1 <- read_excel("Data/1. Odluka.xlsx")
odluka_5 <- read_excel("Data/5. Odluka.xlsx")
odluka_6 <- read_excel("Data/6. Odluka.xlsx")
odluka_7 <- read_excel("Data/7. Odluka.xlsx")
odluka_8 <- read_excel("Data/8. Odluka.xlsx")
odluka_9 <- read_excel("Data/9. Odluka.xlsx")
odluka_10 <- read_excel("Data/10. Odluka.xlsx")

odluka <- bind_rows(odluka_1, odluka_5, 
                    odluka_6, odluka_7, 
                    odluka_8, odluka_9, 
                    odluka_10)

#wrangle the data from "Odluka"
odluka %>% mutate("KC_odluka" = KC) %>% separate_rows(KC, sep = ",") -> odluka
odluka %>% mutate(KC = str_trim(KC)) %>% 
  mutate("KC_2" = str_replace(KC, pattern = 'd.', replacement = "")) -> odluka
```

### The helping data {-}

This is the data that will be joined with the data from "Odluka" files.

```{r}
#data for cadastral units
KO <- read_csv("Data/KO.csv") %>% as_tibble()

#files concerning the cadastral parcels, for whole Croatia
e <- environment()
list.files(path = ".\\KC_zupanije_csv", pattern = ".csv") -> ls_names
str_c(".\\KC_zupanije_csv\\", list.files(path = ".\\KC_zupanije_csv", pattern = ".csv")) -> ls_paths

map(setNames(ls_paths, ls_names), read_csv) %>% as.environment() -> e
e %>% as.list %>% rbindlist(fill = T, idcol = T) -> e_s

#a variable that will be needed later, the KOKC variable
e_s %>% mutate(KOKC = str_c(KO, '_', BROJ_KC)) -> e_s
```

The first joining step: joining the parcels with KO ID number

```{r}
#ID of KO is needed first
odluka %>% left_join(KO, by = c("KO" = "ko_fix", "Sumarija" = "Sumariafix")) -> odluka
odluka$ko_mb <- as.numeric(odluka$ko_mb)
```

Some cadastral units were not joined with the odluka table because the unit names are not typed correctly. The reparation needs to be done by hand. This few units that were left empty in odluka can easily be found.

```{r the problematic data}
#the problematic data needs to be cleaned by hand
#the Na values that are not filled because they were not joined. Why is that so?
odluka %>% filter(is.na(ko_mb)==T) %>% select(KO) %>% unique()
#it is mostly because the names of KO are not matching, they are fixed now

```

# The joining part {-}

```{r}
#join cadastral parcels by ID of cadastral unit, and cadastral parcel number
#before that, joining the ID and number together so that the joining can be faster by checking the double criteria at the same time
odluka %>% mutate(KOKC = str_c(ko_mb, '_', KC_2), 
                  Pov_unique = Povrsina) -> odluka
#the unique area is needed so that the sum is valid later
odluka$Pov_unique[duplicated(odluka$KC_odluka)] <- NA
odluka %>% left_join(e_s, by = c("KOKC" = "KOKC")) -> odluka_spojena

#check of variables
odluka_spojena %>% filter(is.na(.id)==T)
odluka_spojena$Povrsina <- as.numeric(odluka_spojena$Povrsina)
odluka_spojena %>% filter(is.na(Povrsina)==T)
odluka_spojena %>% distinct(KOKC)

#after the check, the rest of the parcels should be joined in second iteration
odluka_spojena %>% filter(is.na(.id)==T) -> odluka_part_2
odluka_part_2 %>% mutate(KOKC = str_c(KOKC, '*')) -> odluka_part_2
#something like this, didn't really work
odluka_part_2 %>% regex_left_join(e_s, by=c("KOKC"="KOKC"))

#split by KO to check the number of KO's
split_list <- split(odluka,odluka$KO)

odluka_spojena %>% write_csv("odluke_sve.csv")

```

## Contracts {-}

For some cadastral parcels the contracts were signed. The "ugovori-MP_fix" table represents those parcels.

```{r}
ugovori_MP_fix <- read_excel("Data/ugovori-MP_fix.xlsx")
ugovori_MP_fix %>% separate_rows(Katastarske_opcine_CONC, 
                                 DIO_CONC, 
                                 POVRSINA_cestica_CONC, 
                                 PRIL_KOORD_CONC, sep = "/")

```

