---
title: "2_data_analysis"
author: "Nikola Gersak"
date: '21 08 2022 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(readxl)

```

## R Markdown

The data concerning cadastral parcels form "the decisions" was wrangled in the first report.
First i'm interested in some descriptive statistical information I can obtain from the data. For example, i might be asking questions such as:

<ol>

which "decisions" carry the most cadastral parcels? 
which USP-s are the most frequent in cadastral parcels?
which sumarijas are the most frequent?
which KO-s are the most frequent? - maybe not
which USP-s gave the most surface of their forests away?
which "decisions" carry the most invalid/non existent cadastral parcels?

what's the  ratio of signed contracts and non signed contracts for cadastral parcels?
which decisions hold the majority of contracts signed?
how is it correlated with years?
how is it correlated with the distance to the sea?
is there correlation with surface and contracts signed?

what are the parcels used now for?
how many of the cadastral parcels were used for the purpose they were intended to?
how many for whom the contracts were signed?
how does the distribution of land use of these parcels look like now?

how many of those parcels fall in the area of natura 2000?
how does the land use of those parcels correlate with the measures for protecting the habitat types of interest?

</ol>


```{r}

#distribution by USP and Sumarija
odluka %>% group_by(USP.x) %>% 
  summarise("Parcels" = n()) %>% 
  ggplot(aes(x = reorder(USP.x, Parcels), 
             y = Parcels, 
             fill = USP.x)) + 
  geom_bar(stat = 'identity') +
  theme_classic()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = c(0.25,0.65))+
  scale_fill_brewer(palette = "Paired")+
  xlab(label = "USP")+
  ylab(label = "Number of parcels")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
#distribution by odluka
odluka %>% group_by(odluka) %>% 
  summarise("Parcels" = n(), 
            odluka_o = as.character(odluka)) %>% 
  unique %>% 
  ggplot(aes(x = reorder(odluka_o, odluka), 
             y = Parcels,
             fill = odluka_o)) + 
  geom_bar(stat = 'identity') +
  theme_classic()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = c(0.25,0.65))+
  scale_fill_brewer(palette = "Set1")+
  xlab(label = "Odluka")+
  ylab(label = "Number of parcels")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#most of surface - chart pie
odluka_spojena %>% group_by(odluka) %>% 
  summarise("Surface" = sum(Povrsina), odluka_o = as.character(odluka)) %>% unique %>% 
  ggplot(aes(x = reorder(odluka_o, odluka), 
             y = Surface,
             fill = odluka_o)) + 
  geom_bar(stat = 'identity') +
  theme_classic()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = c(0.25,0.65))+
  scale_fill_brewer(palette = "Set1")+
  xlab(label = "Odluka")+
  ylab(label = "Area")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

odluka %>% group_by(USP.x) %>% 
  summarise("Surface" = sum(Povrsina), odluka_o = as.character(odluka)) %>% unique %>% 
  ggplot(aes(x = reorder(USP.x, Surface), 
             y = Surface, 
             fill = odluka_o)) + 
  geom_bar(stat = 'identity') +
  theme_classic()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = c(0.25,0.65))+
  scale_fill_brewer(palette = "Paired")+
  xlab(label = "USP")+
  ylab(label = "Area")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("Povrsine u ovisnosti o USP i odlukama.jpg")

```

